A Simple and Strong Baseline for Universal Targeted Attacks on Siamese Visual Tracking

### 2021-05-06 博士/论文提交/adversarial_attack/TCSVT/latex/imperceptible.tex
调整图片位置
按高老师改

Siamese visual tracking is a fundamental research direction in template matching-based tracking besides the correlation filter-based methods. Both of them are aimed to "causally" estimate the positions of a template cropped from the initial video frame in the subsequent frames. Siamese trackers formulate visual tracking as learning cross-correlation similarities between a target template and the candidates in the search region in a convolution fashion. Tracking is then performed by locating the object in the search image region based on the highest visual similarity. This paradigm is formulated as a local one-shot detection task.

@@

\begin{table}
  \centering
  \caption{Overall attack results on the evaluation benchmarks.}
  \begin{tabular}{c c | c | c | c}
  \toprule
  \multirow{2}{*}[-2pt]{Benchmarks} & \multirow{2}{*}[-2pt]{Metrics} & Clean Videos    & \multicolumn{2}{c}{Perturbed Videos}  \\
  \cmidrule{3-5}
                            &                         & Real Traj. & Real Traj. & Fake Traj.     \\ 
  \midrule
  \multirow{2}{*}{OTB-15} 
  & AO   & 0.642 & 0.063 & 0.759\\
  & Precision & 0.861 & 0.092 & 0.795\\
  \midrule
  \multirow{2}{*}{GOT-Val} 
  & SR & 0.897 & 0.123 & 0.890\\
  & AO & 0.760 & 0.153 & 0.840 \\
  \midrule
  \multirow{3}{*}{LaSOT} 
  & Precision  & 0.514 & 0.046 & 0.605\\
  & Norm. Prec.& 0.551 & 0.048 & 0.702\\
  & AO         & 0.525 & 0.069 & 0.691\\
  %& Succ. rate  & 0.626 & 0.016 & 0.834\\
  \midrule
  \multicolumn{2}{c|}{FPS} & 58 & 58 & 58\\
  \bottomrule
  \end{tabular}
  \label{tab:benchmark results}
  \end{table}

@@

In this subsection, we show how to train the video-agnostic perturbations $(\delta, p)$ for Siamese trackers. At beginning, each element in $\delta$ and $p$ is initialized to 0.
During the $k$-th iteration of training, a video $V=\{I_i\}_1^T$ is randomly selected from the training dataset $\mathcal V$. 
Assuming the template perturbation at the $k$-th iteration is $\delta_k \in \mathbb{R}^{127\times 127 \times 3}$, and the adversarial patch is $p_k \in \mathbb{R}^{64\times 64\times 3}$. 
We first randomly pick paired frames $I_t, I_s$ from $V$. The clean template image $\textbf z\in \mathbb{R}^{127\times 127 \times 3}$ is generated according to $I_t$ and $b^{gt}_t$, and the perturbed template image is:
\begin{equation}
\tilde {\textbf z} = \textbf z + \delta_k.
\end{equation}
Similarly, the clean search image $\textbf x \in \mathbb{R}^{303\times 303 \times 3}$ is generated according to $I_s$ and $b^{gt}_s$.
As mentioned before, the patch is regarded as a \textit{fake target} and added into the search images. We force the center position of the \textit{fake target} to near the center position of the real target within a shift range of 64 pixels, where the shift is defined as the maximum range of translation generated from a uniform distribution.
The perturbed search image is generated as follows:
\begin{equation}
\tilde{\textbf x} = A(\textbf x, p_k, \{x_0, y_0, x_1, y_1\}),
\end{equation}
where $ b^{fake} = \{x_0, y_0, x_1, y_1\}$ denote the coordinates of the upper-left and lower-right corners of the fake target on the search image. $A$ is a patch application operator \cite{patch} which adds the patch into the search image $\textbf x$ at location $(\frac{x_0+x_1}{2},\frac{y_0+y_1}{2})$. Subsequently, the SiamFC++ tracker takes $\tilde {\textbf x}$ and $\tilde{\textbf z}$ as input and predicts $\textbf{C, R, Q}$ in an anchor-free manner.

\begin{table}
  \centering
  \caption{Contribution of each loss on GOT-Val.}
  \begin{tabular}{ccc|cc|cc} 
  \toprule
  \multirow{2}{*}[-2pt]{$L_{\text{cls}}$}     & \multirow{2}{*}[-2pt]{$L_{\text{quality}}$} & \multirow{2}{*}[-2pt]{$L_{\text{reg}}$} & \multicolumn{2}{c|}{Fake Traj.}          & \multicolumn{2}{c}{Real Traj.}           \\ 
  \cmidrule{4-7}
                         &                    &                    & AO                    & SR                    & AO                    & SR                     \\ 
  \midrule
  \checkmark   &    &    & 0.643  & 0.726    & 0.282 & 0.274   \\
     & \checkmark   &    & 0.148  & 0.110    & 0.747 & 0.882   \\
     &    & \checkmark   & 0.726  & 0.762    & 0.243 & 0.238   \\
  \checkmark   & \checkmark   & \checkmark   & 0.840  & 0.890    & 0.153 & 0.123   \\ \bottomrule
  \end{tabular}
  \label{tab:loss}
  \end{table}

@@

  \begin{figure}
    \centering
    %\subfigure[Baseline-1]{\label{fig:imperceptiblea}\includegraphics[width=0.11\textwidth]{images_imperceptible/imperceptible/AP.jpg}}
    \subfigure[Baseline-1]{\label{fig:imperceptibleb}\includegraphics[width=0.15\textwidth]{images_imperceptible/imperceptible/UAP.jpg}}
    \subfigure[Baseline-2]{\label{fig:imperceptiblec}\includegraphics[width=0.15\textwidth]{images_imperceptible/imperceptible/AP.jpg}}
    \subfigure[Ours]{\label{fig:imperceptibled}\includegraphics[width=0.15\textwidth]{images_imperceptible/imperceptible/Ours.jpg}}
    \caption{Visualization of perturbations on search images. Baseline-1 generates perturbations added to the whole search image and cannot perform targeted attacks. Baseline-2 pastes an obviously noticeable patch on the search image, which raises the risk of being suspected. Our method adds a small (universal) imperceptible patch to the search image to perform targeted attacks.}
    \label{fig:imperceptible}
  \end{figure}

@@

@@

\begin{table}
  \centering
  \caption{Transferability to different backbones on GOT-Val.}
  \begin{tabular}{c|cc|cc|cc} 
  \toprule
  \multirow{3}{*}[-6pt]{Backbone} & \multicolumn{2}{c|}{Clean Videos}    & \multicolumn{4}{c}{Perturbed Videos}                                        \\ 
  \cmidrule{2-7}
                            & \multicolumn{2}{c|}{Real Traj.} & \multicolumn{2}{c|}{Real Traj.} & \multicolumn{2}{c}{Fake Traj.}  \\ 
  \cmidrule{2-7}
                            & AO    & SR                           & AO    & SR                           & AO    & SR                           \\ 
  \midrule
  GoogLeNet                 & 0.760 & 0.897                        & 0.153 & 0.123                        & 0.840 & 0.890                        \\
  AlexNet                   & 0.720 & 0.850                        & 0.496 & 0.577                        & 0.327 & 0.336                        \\
  ShuffleNet                & 0.766 & 0.888                        & 0.496 & 0.557                        & 0.409 & 0.426                       \\
  \bottomrule
  \end{tabular}
  \label{tab:backbone}
\end{table}

@@

\begin{table}
  \centering
  \caption{Transferability to different tracking architectures on OTB2015.}
  \begin{tabular}{c|cc|cc} 
  \toprule
  \multirow{2}{*}[-2pt]{Trackers} & \multicolumn{2}{c|}{Clean Videos} & \multicolumn{2}{c}{Perturbed Videos}  \\
  \cmidrule{2-5}
                            & AO & Precision              & AO & Precision                   \\
  \midrule
  SiamRPN++                 & 0.676   & 0.879                  & 0.418   & 0.556                       \\
  SiamRPN                   & 0.666   & 0.876                  & 0.483   & 0.643                       \\
  \bottomrule
  \end{tabular}
  \label{tab:arch}
\end{table}

@@

\begin{figure}
  \centering
  \subfigure[64$\times$64]{\label{fig:a}\includegraphics[width=0.15\textwidth]{images_imperceptible/patch_size/64.jpg}}
  \subfigure[32$\times$32]{\label{fig:b}\includegraphics[width=0.15\textwidth]{images_imperceptible/patch_size/32.jpg}}
  \subfigure[16$\times$16]{\label{fig:c}\includegraphics[width=0.15\textwidth]{images_imperceptible/patch_size/16.jpg}}
  \caption{Adversarial patches of different sizes.}
  \label{fig:patch_size_vis}
\end{figure}

@@

We compare the attack method proposed in this paper with the latest attack methods, including the AlexNet-based untargeted attack method CSA \cite{CSA}, and two other targeted attack methods: the AlexNet-based FAN \cite{FAN} and the ResNet50-based TTP \cite{TTP} (Figure \ref{fig:SOTA}). We report the accuracy scores with respect to the real/fake trajectories in table \ref{tab:SOTA}. CSA \cite{CSA} is a Siamese tracking attack method called cooling-shrinking attack. CSA can suppress the peak region on the heat map reflecting the target location, which is used to attack the tracker's targeting ability. CSA can also simultaneously force the tracker's predicted bounding box to shrink, which is used to attack the bounding box regression capability of the tracker. However, this method requires running a generative network for each frame to obtain adversarial information, so the attack time is as high as 4720 ms, making it difficult to meet the demand for real-time tracking. CSA reduces the accuracy score with respect to the real trajectory from 0.851 to 0.458. Due to the limitations of the algorithm design, CSA cannot perform targeted attacks for Siamese trackers. Recently, Liang et al. proposed a fast attack network (FAN) \cite{FAN} for attacking SiamFC trackers. To perform untargeted attacks, FAN proposes a drift loss that shifts the tracker's prediction of the target's position. The tracking error accumulates over time until the tracker loses the target completely. To perform targeted attacks, FAN proposes embedding feature loss for improving the similarity between the features of the adversarial sample and the features of regions specified by a particular trajectory. FAN can reduce the accuracy score with respect to the real trajectory from 0.720 to 0.180. However, the accuracy score with respect to fake trajectories is only 0.420. Similar to SCA, FAN also requires running a generative network for each frame to obtain adversarial information. Nakka et al. \cite{TTP} proposed the temporally transferable perturbations (TTP) for attacking the SiamRPN++ tracker. TTP generates a single adversarial perturbation from the template image and adds this perturbation to each search image of the video. TTP reduces the accuracy score with respect to the real trajectory from 0.910 to 0.080, which is effective for untargeted attacks. However, the accuracy score with respect to the fake trajectory is only 0.692, which demonstrates that TTP's performance of targeted attacks is limited. In addition, this method needs to run a generative network for each video to obtain adversarial information, thus requires the computational and storage resources of the object tracking platform, which makes it difficult to deploy to resource-constrained platforms. Our method can reduce the accuracy score with respect to real trajectories from 0.861 to 0.092. The accuracy score is up to 0.795 with respect to fake trajectories, which is significantly better than other methods, and only needs to perform the add operation to generate adversarial information for any video without gradient optimization or network inference.

>>>>

Siamese visual tracking is a fundamental research direction in template matching-based tracking besides the correlation filter-based methods. Both of them are aimed to ``causally'' estimate the positions of a template cropped from the initial video frame in the subsequent frames. Siamese trackers formulate visual tracking as learning cross-correlation similarities between a target template and the candidates in the search region in a convolution fashion. Tracking is then performed by locating the object in the search image region based on the highest visual similarity. This paradigm is formulated as a local one-shot detection task.

@@

@@

In this subsection, we show how to train the video-agnostic perturbations $(\delta, p)$ for Siamese trackers. At beginning, each element in $\delta$ and $p$ is initialized to 0.
During the $k$-th iteration of training, a video $V=\{I_i\}_1^T$ is randomly selected from the training dataset $\mathcal V$. 
Assuming the template perturbation at the $k$-th iteration is $\delta_k \in \mathbb{R}^{127\times 127 \times 3}$, and the adversarial patch is $p_k$. 
We first randomly pick paired frames $I_t, I_s$ from $V$. The clean template image $\textbf z\in \mathbb{R}^{127\times 127 \times 3}$ is generated according to $I_t$ and $b^{gt}_t$, and the perturbed template image is:
\begin{equation}
\tilde {\textbf z} = \textbf z + \delta_k.
\end{equation}
Similarly, the clean search image $\textbf x \in \mathbb{R}^{303\times 303 \times 3}$ is generated according to $I_s$ and $b^{gt}_s$.
As mentioned before, the patch is regarded as a \textit{fake target} and added into the search images. We force the center position of the \textit{fake target} to near the center position of the real target within a shift range of 64 pixels, where the shift is defined as the maximum range of translation generated from a uniform distribution.
The perturbed search image is generated as follows:
\begin{equation}
\tilde{\textbf x} = A(\textbf x, p_k, \{x_0, y_0, x_1, y_1\}),
\end{equation}
where $ b^{fake} = \{x_0, y_0, x_1, y_1\}$ denote the coordinates of the upper-left and lower-right corners of the fake target on the search image. $A$ is a patch application operator \cite{patch} which adds the patch into the search image $\textbf x$ at location $(\frac{x_0+x_1}{2},\frac{y_0+y_1}{2})$. Subsequently, the SiamFC++ tracker takes $\tilde {\textbf x}$ and $\tilde{\textbf z}$ as input and predicts $\textbf{C, R, Q}$ in an anchor-free manner.

@@

@@

\begin{figure}[t]
  \centering
  %\subfigure[Baseline-1]{\label{fig:imperceptiblea}\includegraphics[width=0.11\textwidth]{images_imperceptible/imperceptible/AP.jpg}}
  \subfigure[Baseline-1]{\label{fig:imperceptibleb}\includegraphics[width=0.15\textwidth]{images_imperceptible/imperceptible/UAP.jpg}}
  \subfigure[Baseline-2]{\label{fig:imperceptiblec}\includegraphics[width=0.15\textwidth]{images_imperceptible/imperceptible/AP.jpg}}
  \subfigure[Ours]{\label{fig:imperceptibled}\includegraphics[width=0.15\textwidth]{images_imperceptible/imperceptible/Ours.jpg}}
  \caption{Visualization of perturbations on search images. Baseline-1 generates perturbations added to the whole search image and cannot perform targeted attacks. Baseline-2 pastes an obviously noticeable patch on the search image, which raises the risk of being suspected. Our method adds a small (universal) imperceptible patch to the search image to perform targeted attacks.}
  \label{fig:imperceptible}
\end{figure}

\begin{table}[t]
  \centering
  \caption{Overall attack results on the evaluation benchmarks.}
  \begin{tabular}{c c | c | c | c}
  \toprule
  \multirow{2}{*}[-2pt]{Benchmarks} & \multirow{2}{*}[-2pt]{Metrics} & Clean Videos    & \multicolumn{2}{c}{Perturbed Videos}  \\
  \cmidrule{3-5}
                            &                         & Real Traj. & Real Traj. & Fake Traj.     \\ 
  \midrule
  \multirow{2}{*}{OTB-15} 
  & AO   & 0.642 & 0.063 & 0.759\\
  & Precision & 0.861 & 0.092 & 0.795\\
  \midrule
  \multirow{2}{*}{GOT-Val} 
  & SR & 0.897 & 0.123 & 0.890\\
  & AO & 0.760 & 0.153 & 0.840 \\
  \midrule
  \multirow{3}{*}{LaSOT} 
  & Precision  & 0.514 & 0.046 & 0.605\\
  & Norm. Prec.& 0.551 & 0.048 & 0.702\\
  & AO         & 0.525 & 0.069 & 0.691\\
  %& Succ. rate  & 0.626 & 0.016 & 0.834\\
  \midrule
  \multicolumn{2}{c|}{FPS} & 58 & 58 & 58\\
  \bottomrule
  \end{tabular}
  \label{tab:benchmark results}
\end{table}

@@

\begin{table}[t]
  \centering
  \caption{Contribution of each loss on GOT-Val.}
  \begin{tabular}{ccc|cc|cc} 
  \toprule
  \multirow{2}{*}[-2pt]{$L_{\text{cls}}$}     & \multirow{2}{*}[-2pt]{$L_{\text{quality}}$} & \multirow{2}{*}[-2pt]{$L_{\text{reg}}$} & \multicolumn{2}{c|}{Fake Traj.}          & \multicolumn{2}{c}{Real Traj.}           \\ 
  \cmidrule{4-7}
                         &                    &                    & AO                    & SR                    & AO                    & SR                     \\ 
  \midrule
  \checkmark   &    &    & 0.643  & 0.726    & 0.282 & 0.274   \\
     & \checkmark   &    & 0.148  & 0.110    & 0.747 & 0.882   \\
     &    & \checkmark   & 0.726  & 0.762    & 0.243 & 0.238   \\
  \checkmark   & \checkmark   & \checkmark   & 0.840  & 0.890    & 0.153 & 0.123   \\ \bottomrule
  \end{tabular}
  \label{tab:loss}
\end{table}

\begin{figure}[t]
  \centering
  \subfigure[64$\times$64]{\label{fig:a}\includegraphics[width=0.15\textwidth]{images_imperceptible/patch_size/64.jpg}}
  \subfigure[32$\times$32]{\label{fig:b}\includegraphics[width=0.15\textwidth]{images_imperceptible/patch_size/32.jpg}}
  \subfigure[16$\times$16]{\label{fig:c}\includegraphics[width=0.15\textwidth]{images_imperceptible/patch_size/16.jpg}}
  \caption{Adversarial patches of different sizes.}
  \label{fig:patch_size_vis}
\end{figure}

\begin{table}[t]
  \centering
  \caption{Transferability to different backbones on GOT-Val.}
  \begin{tabular}{c|cc|cc|cc} 
  \toprule
  \multirow{3}{*}[-6pt]{Backbone} & \multicolumn{2}{c|}{Clean Videos}    & \multicolumn{4}{c}{Perturbed Videos}                                        \\ 
  \cmidrule{2-7}
                            & \multicolumn{2}{c|}{Real Traj.} & \multicolumn{2}{c|}{Real Traj.} & \multicolumn{2}{c}{Fake Traj.}  \\ 
  \cmidrule{2-7}
                            & AO    & SR                           & AO    & SR                           & AO    & SR                           \\ 
  \midrule
  GoogLeNet                 & 0.760 & 0.897                        & 0.153 & 0.123                        & 0.840 & 0.890                        \\
  AlexNet                   & 0.720 & 0.850                        & 0.496 & 0.577                        & 0.327 & 0.336                        \\
  ShuffleNet                & 0.766 & 0.888                        & 0.496 & 0.557                        & 0.409 & 0.426                       \\
  \bottomrule
  \end{tabular}
  \label{tab:backbone}
\end{table}
\begin{table}[t!]
  \centering
  \caption{Transferability to different tracking architectures on OTB2015.}
  \begin{tabular}{c|cc|cc} 
  \toprule
  \multirow{2}{*}[-2pt]{Trackers} & \multicolumn{2}{c|}{Clean Videos} & \multicolumn{2}{c}{Perturbed Videos}  \\
  \cmidrule{2-5}
                            & AO & Precision              & AO & Precision                   \\
  \midrule
  SiamRPN++                 & 0.676   & 0.879                  & 0.418   & 0.556                       \\
  SiamRPN                   & 0.666   & 0.876                  & 0.483   & 0.643                       \\
  \bottomrule
  \end{tabular}
  \label{tab:arch}
\end{table}

@@

@@

@@

We compare the attack method proposed in this paper with the latest attack methods, including the AlexNet-based untargeted attack method CSA \cite{CSA}, and two other targeted attack methods: the AlexNet-based FAN \cite{FAN} and the ResNet50-based TTP \cite{TTP} (Figure \ref{fig:SOTA}). We report the accuracy scores with respect to the real/fake trajectories in table \ref{tab:SOTA}. CSA \cite{CSA} is a Siamese tracking attack method called cooling-shrinking attack. CSA can suppress the peak region on the heat map reflecting the target location, which is used to attack the tracker's targeting ability. CSA can also simultaneously force the tracker's predicted bounding box to shrink, which is used to attack the bounding box regression capability of the tracker. However, this method requires running a generative network for each frame to obtain adversarial information, so the attack time is as high as 4720 ms, making it difficult to meet the demand for real-time tracking. CSA reduces the accuracy score with respect to the real trajectory from 0.851 to 0.458. Due to the limitations of the algorithm design, CSA cannot perform targeted attacks for Siamese trackers. Recently, Liang et al. proposed a fast attack network (FAN) \cite{FAN} for attacking SiamFC trackers. To perform untargeted attacks, FAN proposes a drift loss that shifts the tracker's prediction of the target's position. The tracking error accumulates over time until the tracker loses the target completely. To perform targeted attacks, FAN proposes embedding feature loss for improving the similarity between the features of the adversarial sample and the features of regions specified by a particular trajectory. FAN can reduce the accuracy score with respect to the real trajectory from 0.720 to 0.180. However, the accuracy score with respect to fake trajectories is only 0.420. Similar to SCA, FAN also requires running a generative network for each frame to obtain adversarial information. Nakka et al. \cite{TTP} proposed the temporally transferable perturbations (TTP) for attacking the SiamRPN++ tracker. TTP generates a single adversarial perturbation from the template image and adds this perturbation to each search image of the video. TTP reduces the accuracy score with respect to the real trajectory from 0.910 to 0.080, which is effective for untargeted attacks. However, the accuracy score with respect to the fake trajectory is only 0.692, which demonstrates that TTP's performance of targeted attacks is limited. In addition, this method needs to run a generative network for each video to obtain adversarial information, thus requires the computational and storage resources of the object tracking platform, which makes it difficult to deploy the method to resource-constrained platforms. Our method can reduce the accuracy score with respect to real trajectories from 0.861 to 0.092. The accuracy score is up to 0.795 with respect to fake trajectories, which is significantly better than other methods, and only needs to perform the add operation to generate adversarial information for any video without gradient optimization or network inference.
